---
title: "HW10_Andrew_Goldberg"
author: "Andrew Goldberg"
date: "4/5/2017"
output: html_document
---


####Page 469: #3
The following data were obtained for the growth of a sheep population introduced into a new environment on the island of Tasmania (adapted from J. Davidson, ‘‘On the Growth of the Sheep Population in Tasmania,’’ Trans. R. Soc. S. Australia 62(1938): 342–346).  
  
t(year).....1814.....1824.....1834.....1844....1854.....1864  
p(t)........125......275......830......1200....1750.....1650  

a. Make an estimate of M by graphing P.t/.
```{r 3a}
tYear <- c(1814, 1824, 1834, 1844, 1854, 1864)
pT <- c(125, 275, 830, 1200, 1750, 1650)
plot(tYear, pT)
```
The curve maxes out around 1750, although that could be an outlier since it then declines to 1650. I'd estimate m around 1700. Still, the formula doesn't seem to work with an m below the max, so I'll use 1750. 

b. Plot ln[P/(M-P)] against t. If a logistic curve seems reasonable, estimate rM and t*  
```{r, 3b}
pTB <- log(pT/(1751-pT))
plot(tYear, pTB)
#weird outlier, but still seems reasonable...
rM <- (pTB[6]- pTB[1])/(tYear[6] - tYear[1])
#rM = 0.1071795
C = -4
tStar = -(C/rM)
#t* = 37.32

```

Page 478: #6
Suggest other phenomena for which the model described in the text might be used.  
-fuel usage in motors  
-retail store inventory  
-staffing 
-heating/cooling rooms

####Page 481: #1
a.Using the estimate that db = .054v^2, where 0.054 has dimension ft*hr^2/mi^2, show
that the constant k in Equation (11.29) has the value 19.9 ft/sec2.  
.054*v^2 = v^2/2k  
2*.054 = 1/k  
```{r}
k <- 1/.108
#equating miles to feet and hours to seconds:
k <- k*((5280/3600)^2)
k
```
b. Using the data in Table 4.4, plot db in ft versus v^2/2 in ft2/sec2 to estimate 1/k directly.

```{r}
dbFT <- c(20,28,40.5,52.5, 72, 92.5, 118, 148.5, 182, 220.5, 266, 318, 376)
speedMPH <- c(20,25,30,35,40,45,50,55,60,65,70,75,80)
spMphConv <- ((speedMPH^2)/2) * ((5280/3600)^2)
plot(dbFT, spMphConv)
k2a <- spMphConv[length(spMphConv)]/dbFT[length(dbFT)]
k2a
```



####Page 522: 21 
Oxygen flows through one tube into a liter flask filled with air, and the mixture of oxygen and air (considered well stirred) escapes through another tube. Assuming that air contains 21% oxygen, what percentage of oxygen will the flask contain after 5 L have passed through the intake tube?  
  
Is this a trick question? If 5 litres of oxygen have gone in, and out, then  
V(t) = Vo + (rin - rout)t  
V(t) = .21 + (5-5)t  
V(t) = .21  

####Page 522: 22
If the average person breathes 20 times per minute, exhaling each time 100 in^3 of air containing 4% carbon dioxide. Find the percentage of carbon dioxide in the air of a 10,000-ft^3 closed room 1 hr after a class of 30 students enters. Assume that the air is fresh at the start, that the ventilators admit 1000 ft^3 of fresh air per minute, and that the fresh air contains 0.04% carbon dioxide.
V(t) = Vo + (rin-rout)t  
V(t) = .0004 x (10,000 ft^3) + ((1000ft^3 x .0004)-(30x20x100in^2*.04)
vo = .0004 * 10000 ft^3 = 
```{r}
voFoot <- .0004 * 10000
rinInch <- 20 * 100 * .04 * 30
routFoot <- .0004 * 1000
#V(t) = voFoot + (rinInch - routFoot)t
rInMinOut <- (rinInch*12) - routFoot

```
Didn't have time to finish this :( -- currently on vacation in Asia. 

