---
title: "Homework 15"
author: "Andrew Goldberg"
date: "5/18/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Page B-13: #4 
Discuss how you might go about validating the nuclear arms race model. What data would you collect? Is it possible to obtain the data? 

It's unlikely any of this data would be openly available, but you would need:
1. Amount of missiles or MIRVs/warheads available to each country
2. What the governments consider an unacceptable level of damage (presumably figure out how many missiles would be required to hit the main population and industrial centers of eachs country)
3. Figure out accuracy of each type of missile, along with how easily it would penetrate the opponent country's missile silos, and civil defense
4. There has only been 2 nukes dropped, so there wouldn't be any available sample size for past results, but you could look at the ratios of key world players and their counterparts (usa/russia). 

##Page B-17: #1
1. Build a numerical solution to Equations (15.8).
$$y_{n+1} = 120 + \frac{1}{2}x_{n}$$
$$x_{n+1} = 60 + \frac{1}{3}y_{n}$$
$$x_{0}=100$$
$$y_{0}=200$$

```{r}
y17 <- function(x){120 + .5 * x}
x17 <- function(y){60 + (1/3)*y}

q17 <- function(x0, y0, iters){
  n <- c(0:iters)
  x <- c(x0, 1:iters)
  y <- c(y0, 1:iters)
  
  for(i in 1:iters){
    x[i+1] <- x17(y[i])
    y[i+1] <- y17(x[i])
  }
  solution <- data.frame(n,x,y)
  return(solution)
}

q17(100,200,17)
q17dat <- q17(100,200,17)

#A: graph results
plot(q17dat$n, q17dat$y, type="l", col="green", ylim=(c(0,200)))
lines(q17dat$n, q17dat$x, type="l", col="red")

#B: Is an equlibrium value reached?
#An equilibrium value is reached at x=120, y=180

#C: Try other starting values. Do you think the equilibrium value is stable?
q17(200,100,17)[15:17,] #looks stable
q17(150,150,17)[15:17,] #still looks stable

#D: Explore other values for the survival coefficients of Countries X and Y . Describe your results.
y17d <- function(sc, x){120 + sc * x}
x17d <- function(sc, y){60 + sc * y}

q17d <- function(x0, y0, iters, scx, scy){
  n <- c(0:iters)
  x <- c(x0, 1:iters)
  y <- c(y0, 1:iters)
  
  for(i in 1:iters){
    x[i+1] <- x17d(scx, y[i])
    y[i+1] <- y17d(scy, x[i])
  }
  solution <- data.frame(n,x,y)
  return(solution)
}

#at .7/.4 (all three reach equilibrium at 200/200 ish)
q17d(100,200,17, .7, .4)[15:17,] 
q17d(200,100,17, .7, .4)[15:17,] 
q17d(150,150,17, .7, .4)[15:17,] 

#at .9/.1 (all three reach equilibrium at 184.6/138.4 ish)
q17d(100,200,17, .9, .1)[15:17,] 
q17d(200,100,17, .9, .1)[15:17,] 
q17d(150,150,17, .9, .1)[15:17,] 

#So takeaway is that this is a generally stable system, and the survival coefficient defines the equlibrium values
```

##Page B-21: #4
Verify the result that the marginal revenue of the $q+1st$ unit equals the price of that unit minus the loss in revenue on previous units resulting from price reduction.

I don't fully understand this question because the preceding explanation of the 15.3 model of the market mentions in two separate paragraphs that the price doesn't fluctuate in a large market and revenue stays linear no matter how much product is produced...

If its a smaller market and price then fluctutes with quantity, then yes, it makes inherent sense that as total revenue increases due to increased production, then the lowered price of the most marginal unit would apply to all units on sale. 
$$MR(q) = TR(q+1) - TR(q)$$
$$TR(q+1) = price(q+1)$$
With standard revenue as described in the chapter:
$$TR(q) = price(q) * quantity(q)$$

But with a lowered price:
$$TR(q) = (price(q+1)*quantity) - (price(q - (q+1)))*quantity)$$
So..
$$MR(q) = TR(q+1) - (price(q+1)*quantity) - (price(q - (q+1)))*quantity)$$

##Page B-25: #1
Show that when the demand curve is very steep, a tax added to each item sold will fall primarily on consumers. 

Setting blue to the demand curve, green to the supply curve, black denoting quantify shifting, and red to the amount of tax...we see that there is a 6 dollar price increase on consumers, with only about 1 dollar of the tax paid by the company. 
```{r}
demand <- function(a,b,p){a-(b*p)}
supply <- function(a,b,p){a+(b*p)}

plot(demand(50,50,seq(0,50,.05)), ylim = c(0,50), xlim = c(0,100), col = "blue")
lines(supply(-2,5, seq(0,50,.05)), col = "green")
lines(supply(5,5, seq(0,50,.05)), col = "green")
abline(h = 10, col = "red")
abline(h = 9, col = "red")
abline(h = 3, col = "red")
abline(v = 20)
abline(v = 17)
```

Now show that when the demand curve is more nearly horizontal, the tax is paid mostly by the industry. 
```{r}
demand <- function(a,b,p){a-(b*p)}
supply <- function(a,b,p){a+(b*p)}

plot(demand(20,1,seq(0,50,.05)), ylim = c(0,50), xlim = c(0,100), col = "blue")
lines(supply(-2,5, seq(0,50,.05)), col = "green")
lines(supply(5,5, seq(0,50,.05)), col = "green")
abline(h = 23.5, col = "red")
abline(h = 18, col = "red")
abline(h = 16.6, col = "red")
abline(v = 53)
abline(v = 75)
```
What if the supply curve is very steep? 

Consumers pay far less of the tax:
```{r}
demand <- function(a,b,p){a-(b*p)}
supply <- function(a,b,p){a+(b*p)}

plot(demand(40,10,seq(0,50,.05)), ylim = c(0,50), xlim = c(0,100), col = "blue")
lines(supply(-5,30, seq(0,50,.05)), col = "green")
lines(supply(5,30, seq(0,50,.05)), col = "green")
abline(h = 40, col = "red")
abline(h = 31, col = "red")
abline(h = 29, col = "red")
abline(v = 18.5)
abline(v = 24)
```
What if the supply curve is nearly horizontal?

Most of the tax is passed onto consumers
```{r}
demand <- function(a,b,p){a-(b*p)}
supply <- function(a,b,p){a+(b*p)}

plot(demand(40,10,seq(0,50,.05)), ylim = c(0,50), xlim = c(0,100), col = "blue")
lines(supply(20,.7, seq(0,50,.05)), col = "green")
lines(supply(10,.7, seq(0,50,.05)), col = "green")
abline(h = 22, col = "red")
abline(h = 21, col = "red")
abline(h = 12, col = "red")
abline(v = 39)
abline(v = 57)
```
##Page B-29: #1
Consider the graphical model in Figure 15.27. Argue that if the demand curve fails to shift significantly to the left, an increase in the equilibrium quantity could occur after the crisis.

If demand doesn't decrease enough with both more efficient transportation and more product on the market, then it's likely that the demand is still quite strong with a steep curve, which would lead to the gas companies continuing the raise the cost and pay a smaller percentage of the taxes. 
