---
title: "Assignment 13"
author: "Andrew Goldberg"
date: "December 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r q1}
#Write a program to compute the derivative of f(x) = x^3 + 2x^2 at any value of x.

first.derivative <- function(x){
  h <- .000000001 #limit as h approaches 0
  (((x+h)^3) + (2*((x+h)^2)) - (x^3) - (2*(x^2)))/h
}
first.derivative(3)

#Now, write a program to compute the area under the curve for the function 3x^2+4x in the range x = [1,3]. You should first split the range into many small intervals using some really small change x value (say 1e-6) and then compute the approximation to the area under the curve.

second.derivative <-function(x1, x2)
  {
  second.integral <- function(x){((3*(x^3))/3 + ((4*(x^2))/2))} #set integral function
  total.area <- 0
  total.intervals <- ((x2 - x1)/.000001) #calculate amount of intervals to run through
  spot <- x1
  for(i in 1:total.intervals){
    interval.area <- second.integral(spot + .000001) - second.integral(spot) #area at each interval
    total.area <- total.area + interval.area #accumulate interval areas
    spot <- spot + .000001
    }
  print(total.area)
  }

second.derivative(1,3)
```

Solving Analytically
integral[f(x)g'(x)dx] = f(x)g(x) - integral [f'(x)g(x)dx]

A. integral[sin(x)cos(x)dx
assignments: f(x) = sin(x), f'(x) = cos(x), g(x) = sin(x), g'(x) = cos(x)
integral[sin(x)cos(x)dx] = sin(x)sin(x) - integral[cos(x)sin(x)dx]

then..integral[cos(x)sin(x)dx]
assignments: 
f(x) = cos(x), f'(x) = -sin(x), g(x) = -cos(x), g'(x) = sin(x)
integral[cos(x)sin(x)dx] = cos(x)-cos(x) - integral[-sin(x)-cos(x)dx]

plugging in to original...
integral[sin(x)cos(x)dx] = sin(x)sin(x) - cos(x)cos(x) + integral[-sin(x)-cos(x)dx]
integral[sin(x)cos(x)dx] = sin(x)sin(x) - cos(x)cos(x) - integral[sin(x)cos(x)dx]
2(integral[sin(x)cos(x)dx]) = sin(x)sin(x) - cos(x)cos(x)
integral[sin(x)cos(x)dx] = (sin(x)sin(x) - cos(x)cos(x))/2 + C

...a triganomic identity is cos(a+b) = cos(a)cos(b) - sin(a)sin(b), so...
integral[sin(x)cos(x)dx] = (-cos(2x))/2 + C

B. integral[x^2e^xdx
assignments: f(x) = x^2, f'(x) = 2x, g(x) = e^x, g'(x) = e^x

integral[x^2e^xdx = x^2e^x - 2(integral[x(e^x)dx)

integral[x(e^x)dx
assignments: f(x) = x, f'(x) = 1, g(x) = e^x, g'(x) = e^x
integral[x(e^x)dx = xe^x - integral[e^xdx]
integral[x(e^x)dx = xe^x - e^x
plugging in...
integral[x^2e^xdx = x^2e^x - 2(xe^x - e^x) + C
integral[x^2e^xdx = x^2(e^x) -2(x(e^x)) + 2e^x + C

C. d/dx(x(cos(x)))
= cos(x)-sin(x)(x)

D. d/dx(e^x^4)
= 4x(e^x^4)

