---
title: "Inference for numerical data"
output:
  pdf_document: default
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
---

## North Carolina births

In 2004, the state of North Carolina released a large data set containing 
information on births recorded in this state. This data set is useful to 
researchers studying the relation between habits and practices of expectant 
mothers and the birth of their children. We will work with a random sample of 
observations from this data set.

## Exploratory analysis

Load the `nc` data set into our workspace.

```{r load-data}
load("C:/Users/Andrew/Documents/R/win-library/3.2/IS606/labs/Lab5/more/nc.RData")
getwd()
```

We have observations on 13 different variables, some categorical and some 
numerical. The meaning of each variable is as follows.

variable         | description
---------------- | -----------
`fage`           | father's age in years.
`mage`           | mother's age in years.
`mature`         | maturity status of mother.
`weeks`          | length of pregnancy in weeks.
`premie`         | whether the birth was classified as premature (premie) or full-term.
`visits`         | number of hospital visits during pregnancy.
`marital`        | whether mother is `married` or `not married` at birth.
`gained`         | weight gained by mother during pregnancy in pounds.
`weight`         | weight of the baby at birth in pounds.
`lowbirthweight` | whether baby was classified as low birthweight (`low`) or not (`not low`).
`gender`         | gender of the baby, `female` or `male`.
`habit`          | status of the mother as a `nonsmoker` or a `smoker`.
`whitemom`       | whether mom is `white` or `not white`.

1.  What are the cases in this data set? How many cases are there in our sample?
The cases are births recorded in North Carolina, no apparant date specified

As a first step in the analysis, we should consider summaries of the data. This
can be done using the `summary` command:

```{r summary, eval=FALSE}
summary(nc)
```

As you review the variable summaries, consider which variables are categorical 
and which are numerical. For numerical variables, are there outliers? If you 
aren't sure or want to take a closer look at the data, make a graph.

Consider the possible relationship between a mother's smoking habit and the 
weight of her baby. Plotting the data is a useful first step because it helps 
us quickly visualize trends, identify strong associations, and develop research
questions.

2.  Make a side-by-side boxplot of `habit` and `weight`. What does the plot 
highlight about the relationship between these two variables?

```{r boxplots-habit-weight}
boxplot(weight~habit, data = nc)
```
The means are similar and non smokers have a greater iqr stretching into heavier babies. Non smokers also have more lighter babies, which is probably a function of having a higher proportion of non smokers. 

The box plots show how the medians of the two distributions compare, but we can
also compare the means of the distributions using the following function to 
split the `weight` variable into the `habit` groups, then take the mean of each
using the `mean` function.

```{r by-means, eval=FALSE}
by(nc$weight, nc$habit, mean)
```

There is an observed difference, but is this difference statistically 
significant? In order to answer this question we will conduct a hypothesis test
.

## Inference

3.  Check if the conditions necessary for inference are satisfied. Note that 
you will need to obtain sample sizes to check the conditions. You can compute 
the group size using the same `by` command above but replacing `mean` with 
`length`.

```{r}
library(ggplot2)
#Presumably the sample is less than 10% of the population
summary(nc$habit)
#each variable has more than 30 observations
ggplot(nc, aes(weight)) + geom_bar(binwidth = .5) + facet_wrap(~ habit) 
#data follows fairly normal distribution, slight left skew
```

4.  Write the hypotheses for testing if the average weights of babies born to 
smoking and non-smoking mothers are different.

Ho = Average baby weights are the same among both smoking and non-smoking mothers
Ha = Average baby weights differ between smoking and non-smoking mothers

Next, we introduce a new function, `inference`, that we will use for conducting
hypothesis tests and constructing confidence intervals. 

```{r inf-weight-habit-ht, eval=FALSE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

Let's pause for a moment to go through the arguments of this custom function. 
The first argument is `y`, which is the response variable that we are 
interested in: `nc$weight`. The second argument is the explanatory variable, 
`x`, which is the variable that splits the data into two groups, smokers and 
non-smokers: `nc$habit`. The third argument, `est`, is the parameter we're 
interested in: `"mean"` (other options are `"median"`, or `"proportion"`.) Next
we decide on the `type` of inference we want: a hypothesis test (`"ht"`) or a 
confidence interval (`"ci"`). When performing a hypothesis test, we also need 
to supply the `null` value, which in this case is `0`, since the null 
hypothesis sets the two population means equal to each other. The `alternative`
hypothesis can be `"less"`, `"greater"`, or `"twosided"`. Lastly, the `method` 
of inference can be `"theoretical"` or `"simulation"` based.

5.  Change the `type` argument to `"ci"` to construct and record a confidence 
interval for the difference between the weights of babies born to smoking and 
non-smoking mothers.

```{r inf-weight-habit-ci}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")
```

By default the function reports an interval for ($\mu_{nonsmoker} - \mu_{smoker}$)
. We can easily change this order by using the `order` argument:

```{r inf-weight-habit, eval=FALSE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("smoker","nonsmoker"))
```

* * *

## On your own

-   Calculate a 95% confidence interval for the average length of pregnancies 
(`weeks`) and interpret it in context. Note that since you're doing inference 
on a single population parameter, there is no explanatory variable, so you can 
omit the `x` variable from the function.

```{r}
inference(y=nc$weeks, est = "mean", type = "ci", null = 0, alternative = "twosided", method = "theoretical")
```
The average length of pregnancies is 38 weeks, with a confidence interval of (38.2 weeks to 38.5 weeks) at the 95% confidence level. 

-   Calculate a new confidence interval for the same parameter at the 90% 
confidence level. You can change the confidence level by adding a new argument 
to the function: `conflevel = 0.90`.

```{r}
inference(y=nc$weeks, est = "mean", type = "ci", null = 0, alternative = "twosided", method = "theoretical", conflevel = 0.90)
```
(38.182, 38.4873)

-   Conduct a hypothesis test evaluating whether the average weight gained by 
younger mothers is different than the average weight gained by mature mothers.

```{r}
inference(y=nc$weight, x=nc$mature, est = "mean", type = "ht", null = 0, alternative = "twosided", method = "theoretical", conflevel = 0.95)
```
P-value is .8526 which is larger than .05, so we will fail to reject the null

-   Now, a non-inference task: Determine the age cutoff for younger and mature 
mothers. Use a method of your choice, and explain how your method works.

```{r}
#plot mother age by maturity, looks like they split around 34/35
plot(nc$mage, nc$mature)

#checking min age for mature mothers shows 35
min((nc[nc$mature == "mature mom",])$mage)

#and checking max age for young mothers is 34
max((nc[nc$mature == "younger mom",])$mage)
```

-   Pick a pair of numerical and categorical variables and come up with a 
research question evaluating the relationship between these variables. 
Formulate the question in a way that it can be answered using a hypothesis test
and/or a confidence interval. Answer your question using the `inference` 
function, report the statistical results, and also provide an explanation in 
plain language.

Is there a difference in the amount of weight gained between white and non-white mothers? 
Ho: There is no difference between the weight gained by white and non-white mothers during pregnancy. (The difference of averages = 0)

Ha: There is a difference between the amount of weight gained between white and non-white mothers during pregnancy (The difference of averages does not equal 0)

```{r}
inference(y=nc$gained, x=nc$whitemom, est = "mean", type = "ht", null = 0, alternative = "twosided", method = "theoretical", conflevel = 0.95)
```

White mothers have mean weight gain of 31 units (pounds?), while non-White mothers have average weight gain of 28.7 units. Using a hypothesis test with confidence at the 95% level, the p-value is .0286, so we would reject the null in favor of the alternative. In this case, that White mothers gain more weight than non-white mothers. 

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was adapted for OpenIntro by Mine &Ccedil;etinkaya-Rundel from a lab 
written by the faculty and TAs of UCLA Statistics.
</div>
